---
author: "By Forrest W. Crawford and Whitespace Solutions"
output: html_document
params:
  town: "New Haven"
---

```{r echo=FALSE, messages=FALSE}
library(leaflet) 
library(tigris)
library(lubridate)
library(stringr)
library(htmltools)
library(htmlwidgets)
library(sf)

load("ct_data.Rsave")

town = params$town

alltowns = unique(ct_towns$NAME)
if(!(town %in% alltowns)) {stop(paste("Town", town, "not found."))}

ct_towns = ct_towns %>% sf::st_transform(4326)
ct_cbgs = ct_cbgs %>% sf::st_transform(4326)

# should be prior week using today() 
startdate = min(dat_by_town$date)
enddate = max(dat_by_town$date)






town_boundary = ct_towns[ct_towns$NAME==town,]
suppressWarnings(suppressMessages(town_boundary_buf <- st_buffer(town_boundary, dist=0.005)))
suppressMessages(cbgs_in_town <- st_within(ct_cbgs, town_boundary_buf) %>% lengths > 0 )
town_cbgs = ct_cbgs[cbgs_in_town,]





dat_town = dat_by_town[dat_by_town$town == town,]

lastdate = max(dat_by_town$date)
oneweekago =  lastdate - days(6)
# these are the same length
lastweek = interval(lastdate - days(6), lastdate)
lastweek2 = interval(lastdate - days(13), lastdate - days(7))


dat_town_lastweek  = dat_town[dat_town$date %within% lastweek,]
dat_town_lastweek2 = dat_town[dat_town$date %within% lastweek2,]




##########
# copied from app 

dat_by_town_lastweek_mean = aggregate(prob_sum_contact ~ town, data=dat_by_town_lastweek, mean)
dat_by_town_lastweek2_mean = aggregate(prob_sum_contact ~ town, data=dat_by_town_lastweek2, mean)

ntowns = dim(dat_by_town_lastweek_mean)[1]

dat_by_town_lastweek_mean$rank = ntowns + 1 - rank(dat_by_town_lastweek_mean$prob_sum_contact)
dat_by_town_lastweek2_mean$rank = ntowns + 1 - rank(dat_by_town_lastweek2_mean$prob_sum_contact)

dat_by_town_week_comparison = merge(dat_by_town_lastweek_mean, dat_by_town_lastweek2_mean, by="town", suffixes=c("_lastweek", "_lastweek2"))

rownames(dat_by_town_week_comparison) <- NULL

ord = order(dat_by_town_week_comparison$prob_sum_contact_lastweek, decreasing=TRUE)

dat_by_town_week_comparison = dat_by_town_week_comparison[ord,]

names(dat_by_town_week_comparison) <- c("Town", "Contact this week", "Rank this week", "Contact last week", "Rank last week")

ntowns = dim(dat_by_town_week_comparison)[1]

applyQuartiles <- function(x) {
  cut(x, breaks=c(quantile(x, probs = seq(0, 1, by = 0.25))), 
      labels=c(
      "excellent, in the first quartile",
      "good, in the second quartile",
      "poor, in the third quartile",
      "severe, in the top quartile"), include.lowest=TRUE)
}

# severity for all towns
dat_by_town_week_comparison$town_severity = applyQuartiles(dat_by_town_week_comparison$`Contact this week`)

# this town
town_severity = dat_by_town_week_comparison$town_severity[dat_by_town_week_comparison$Town == town]

town_rank_lastweek  = dat_by_town_lastweek_mean$rank[dat_by_town_lastweek_mean$town == town]
town_rank_lastweek2  = dat_by_town_lastweek2_mean$rank[dat_by_town_lastweek2_mean$town == town]
town_rank_change = town_rank_lastweek - town_rank_lastweek2
town_rank_change_string = paste(ifelse(town_rank_change>0,"+",ifelse(town_rank_change<0,"-","")), town_rank_change, sep="")

town_contact_lastweek = dat_by_town_lastweek_mean$prob_sum_contact[dat_by_town_lastweek_mean$town == town]
town_contact_lastweek2 = dat_by_town_lastweek2_mean$prob_sum_contact[dat_by_town_lastweek2_mean$town == town]

town_trend = (town_contact_lastweek - town_contact_lastweek2)/town_contact_lastweek2
town_trend_string = paste(ifelse(town_trend>0,"+",ifelse(town_trend<0,"-","")), format(town_trend,digits=2), "%", sep="")



# end copied from app
##################



dat_cbg = dat_by_cbg[dat_by_cbg$geoid %in% town_cbgs$GEOID,]

dat_cbg_lastweek  = dat_cbg[dat_cbg$date %within% lastweek,]
dat_cbg_lastweek2 = dat_cbg[dat_cbg$date %within% lastweek2,]


#dat_cbg_lastweek  = dat_cbg[dat_cbg$date >= startdate & dat_cbg$date <= enddate,]

town_lastweek_mean = aggregate(prob_sum_contact ~ town, data=dat_town_lastweek, mean)
town_lastweek2_mean = aggregate(prob_sum_contact ~ town, data=dat_town_lastweek2, mean)

cbg_lastweek_means = aggregate(prob_sum_contact ~ geoid, data=dat_cbg_lastweek, mean)
cbg_lastweek2_means = aggregate(prob_sum_contact ~ geoid, data=dat_cbg_lastweek2, mean)

cbg_lastweek_means_home = aggregate(prob_sum_home ~ geoid, data=dat_cbg_lastweek, mean)

cbg_historical = merge(cbg_lastweek_means, cbg_lastweek2_means, by="geoid", suffixes=c("_lastweek", "_lastweek2"))

cbg_historical = merge(cbg_historical, cbg_lastweek_means_home, by="geoid")

cbg_historical$change = cbg_historical$prob_sum_contact_lastweek - cbg_historical$prob_sum_contact_lastweek2
cbg_historical$pchange = cbg_historical$change / cbg_historical$prob_sum_contact_lastweek2
cbg_historical$townrank1 = length(cbg_historical$change) + 1 - rank(cbg_historical$prob_sum_contact_lastweek)
cbg_historical$townrank2 = length(cbg_historical$change) + 1 - rank(cbg_historical$prob_sum_contact_lastweek2)

ord = order(cbg_historical$townrank1, decreasing=FALSE)

cbg_historical = cbg_historical[ord,]


dat_merged_cbgs_historical = geo_join(town_cbgs, cbg_historical, by_sp="GEOID", by_df="geoid", how="left") %>% 
                    sf::st_transform(4326)



##################33

# compute view bounds
bdry = as.numeric(st_bbox(town_boundary))
```

---
title: `r town`, Connecticut | Social Contact Report (beta)
subtitle: `r format(oneweekago, "%B %d")` -- `r format(lastdate, "%B %d, %Y")`  
---

This report provides data and visualizations showing trends in social contact patterns in Connecticut.  When combined with data on the geographic patterns of testing and cases, this information may assist public health officials and policymakers in taking actions to mitigate the spread of COVID-19 in Connecticut.  High contact does not necessarily mean that COVID-19 transmission is happening, but close interpersonal contact is a necessary condition for transmission to occur.  Statewide interactive maps are available from the [interactive web application](https://forrestcrawford.shinyapps.io/ct_social_distancing/).

### Executive Summary 

Overall the contact rate in `r town` is `r town_severity` of all `r ntowns` towns in Connecticut this week.  Contact in `r town` is ranked number `r town_rank_lastweek` in Connecticut this week, of `r ntowns` towns, which means that `r ifelse(town_rank_lastweek<50,"only","")` `r town_rank_lastweek-1` towns in Connecticut had more contact than `r town`.  Last week `r town` was ranked `r town_rank_lastweek2`, for a change of `r town_rank_change_string`.  Contact in `r town` is trending `r ifelse(town_trend<0,"downward", "upward")`, with a `r town_trend_string` change from last week. 


<div class = "row">
<div class = "col-md-6">

#### Areas where contact is happening

This map shows <font color=\"#00AA00\"><b>contact rates</b></font> in `r town` census block groups.  Dark green regions are areas where communications, policies, or guidelines to *improve social distancing* may be helpful.  

```{r echo=FALSE, out.width='100%'}


pal_cbg_contact <- colorNumeric("Greens", domain=dat_merged_cbgs_historical$prob_sum_contact_lastweek) 


pal_points_of_interest = colorFactor(palette = c("blue", "orange", "purple", "red", "yellow"), 
                                     levels = c("Hospital", "House of Worship", "Sports Venue", "Mall", "University"))

pointrad = 5

basemap = leaflet() %>% 
          addLayersControl(
            position = "topright",
            baseGroups = c("CartoDB.Positron", "Esri.WorldStreetMap", "CartoDB.Voyager", "Esri.WorldImagery"),
            overlayGroups = c("Metric", "Points of Interest"),
            options = layersControlOptions(collapsed = TRUE)) %>% 
          hideGroup(c("Points of Interest")) %>%
          addProviderTiles("CartoDB.Positron", group="CartoDB.Positron") %>% 
          addProviderTiles("Esri.WorldImagery", group="Esri.WorldImagery") %>%
          addProviderTiles("Esri.WorldStreetMap", group="Esri.WorldStreetMap") %>%
          addProviderTiles("CartoDB.Voyager", group="CartoDB.Voyager") %>%
          addPolygons(data=town_boundary, fill=NULL, stroke=TRUE, weight=8, fillOpacity=0, color="black") 

map = basemap %>% addPolygons(data=dat_merged_cbgs_historical, 
                      stroke = TRUE, weight=1, col="gray", smoothFactor = 0.1, fillOpacity = 0.8, 
                      fillColor = ~pal_cbg_contact(dat_merged_cbgs_historical$prob_sum_contact_lastweek), 
                      group = "Metric",
                      label = sprintf("<strong>%s</strong><br/>%g", dat_merged_cbgs_historical$GEOID, dat_merged_cbgs_historical$prob_sum_contact_lastweek) %>% lapply(htmltools::HTML),
                              labelOptions = labelOptions(
                                style = list("font-weight" = "normal", padding = "3px 8px", "color" = "green"),
                                textsize = "15px", direction = "auto"))  %>%
             fitBounds(bdry[1], bdry[2], bdry[3], bdry[4]) %>%
              addCircleMarkers(data=points_of_interest, lat = ~ lat, lng = ~ lon, 
                 weight = 1, stroke=FALSE, radius = pointrad, fillOpacity = 0.3, 
                 group = "Points of Interest",
                 color = ~pal_points_of_interest(type),
                 label = sprintf("<strong>%s: %s</strong><br/>%s %s", points_of_interest$type, points_of_interest$name, points_of_interest$address, points_of_interest$city) %>% lapply(htmltools::HTML),
                 labelOptions = labelOptions(
                   style = list("font-weight" = "normal", padding = "3px 8px"),
                   textsize = "15px", direction = "auto")) %>% 
              addLegend("bottomright", pal = pal_cbg_contact, values = dat_merged_cbgs_historical$prob_sum_contact_lastweek, title = "Contact") 

map
```
</div>

<div class = "col-md-6">

#### Areas where contacting individuals reside

This map shows <font color=\"#0000AA\"><b>home residence locations</b></font> of individuals having contact in `r town` census block groups.  Dark blue regions are areas where enhanced *testing and contact tracing* may be helpful. 


```{r echo=FALSE, out.width='100%'}

pal_cbg_home = colorNumeric("Blues", domain=dat_merged_cbgs_historical$prob_sum_home) 

map2 = basemap %>% addPolygons(data=dat_merged_cbgs_historical, 
                      stroke = TRUE, weight=1, col="gray", smoothFactor = 0.1, fillOpacity = 0.8, 
                      fillColor = ~pal_cbg_home(dat_merged_cbgs_historical$prob_sum_home), 
                      group = "Metric",
                      label = sprintf("<strong>%s</strong><br/>%g", dat_merged_cbgs_historical$GEOID, dat_merged_cbgs_historical$prob_sum_home) %>% lapply(htmltools::HTML),
                              labelOptions = labelOptions(
                                style = list("font-weight" = "normal", padding = "3px 8px", "color" = "blue"),
                                textsize = "15px", direction = "auto"))  %>%
               fitBounds(bdry[1], bdry[2], bdry[3], bdry[4]) %>%
                             addCircleMarkers(data=points_of_interest, lat = ~ lat, lng = ~ lon, 
                 weight = 1, stroke=FALSE, radius = pointrad, fillOpacity = 0.3, 
                 group = "Points of Interest",
                 color = ~pal_points_of_interest(type),
                 label = sprintf("<strong>%s: %s</strong><br/>%s %s", points_of_interest$type, points_of_interest$name, points_of_interest$address, points_of_interest$city) %>% lapply(htmltools::HTML),
                 labelOptions = labelOptions(
                   style = list("font-weight" = "normal", padding = "3px 8px"),
                   textsize = "15px", direction = "auto")) %>% 
   addLegend("bottomright", pal = pal_cbg_home, values = dat_merged_cbgs_historical$prob_sum_home, title = "Home") 

map2
```
</div>
</div>

Alternative base maps and points of interest can be displayed via the layers menu in the upper right corner.

The plots below show trends in `r town` over time in <font color=\"#00AA00\"><b>contact rates</b></font>, with trends for individual census block groups shown in below. The five block groups with the top contact rates this week are shown in black. Gray bars represent weekends. 

```{r, echo=FALSE, out.width='100%', fig.width=12, fig.height=6}

saturdays = seq(mdy("2/1/2020"), today(), by="week")
sundays = seq(mdy("2/2/2020"), today(), by="week")



contact_max = max(dat_town$prob_sum_contact)

par(mar=c(2.5,4,1,1), mfrow=c(2,1))

# town total

plot(dat_town$date, dat_town$prob_sum_contact, type="l", col="green", lwd=3, xlim=c(min(dat_town$date)-1, max(dat_town$date)+14),
         ylim=c(0,max(dat_town$prob_sum_contact)*1.1), xlab="", ylab="Contacts", axes=FALSE, main=paste(town,"total"))
dateseq = c(seq(min(dat_town$date), max(dat_town$date), by="month"), max(dat_town$date))
axis(1,at=dateseq, lab=format(dateseq, "%b %d"))
axis(2)
for(i in 1:length(saturdays)) { rect(saturdays[i]-1/2, 0, sundays[i]+1/2, 5e5, col=rgb(0,0,0,alpha=0.1), border=NA) }

# CBGs

plot(0, type="n", xlim=c(min(dat_cbg$date)-1, max(dat_cbg$date)+14),
     ylim=c(0,max(dat_cbg$prob_sum_contact)*1.1),
     xlab="", 
     ylab="Contacts", axes=FALSE, main="Top block groups")
axis(1,at=dateseq, lab=format(dateseq, "%b %d"))
axis(2)


for(cbg_id in dat_cbg$geoid) {
  cbg_rank = cbg_historical$townrank1[cbg_historical$geoid == cbg_id]
  cbg_dates = dat_cbg$date[dat_cbg$geoid == cbg_id]
  cbg_contact = dat_cbg$prob_sum_contact[dat_cbg$geoid == cbg_id]
  cbg_col = ifelse(cbg_rank<=5,"black", "gray")
  lines(cbg_dates, cbg_contact,  col=cbg_col)
  text(cbg_dates[length(cbg_dates)], cbg_contact[length(cbg_contact)], cbg_id, col=cbg_col, pos=4)
}

for(i in 1:length(saturdays)) { rect(saturdays[i]-1/2, 0, sundays[i]+1/2, 5e5, col=rgb(0,0,0,alpha=0.1), border=NA) }

```


### Top contact areas in `r town` 

The table below shows contact last week by census block groups in `r town`, along with points of interest (hospitals, houses of worship, malls, and sports venues) that may contribute to contact rates.  Also shown are percent change and rank in `r town`. Find specific block groups or points of interest in the search box. The table can be sorted by value for each column. 

```{r echo=FALSE}
library(formattable)
library(DT)

# see here: https://www.displayr.com/formattable/

rownames(dat_merged_cbgs_historical) <- NULL
options(scipen=999) # don't use sci notation for numbers
#cbg_historical$change = signif(cbg_historical$change, digits=2)


#dat_merged_cbgs_historical$poi = str_replace_all(dat_merged_cbgs_historical$poi, "^\\| ", "")
#dat_merged_cbgs_historical$poi = str_replace_all(dat_merged_cbgs_historical$poi, "\\|", "/")


dat_merged_cbgs_historical$`Geo ID` = dat_merged_cbgs_historical$GEOID
dat_merged_cbgs_historical$`Points of interest` = dat_merged_cbgs_historical$poi
dat_merged_cbgs_historical$`Contact this week` = dat_merged_cbgs_historical$prob_sum_contact_lastweek
dat_merged_cbgs_historical$`% change` = percent(dat_merged_cbgs_historical$pchange)
dat_merged_cbgs_historical$`Town rank` = dat_merged_cbgs_historical$townrank1

ord = order(dat_merged_cbgs_historical$townrank1)
dat_merged_cbgs_historical = dat_merged_cbgs_historical[ord,]


as.datatable(
  formattable(dat_merged_cbgs_historical, list(
                         #`Geo ID` = formatter("span", style = ~ style(color = "black",font.weight = "bold")),
                         #`Points of Interest` = formatter("span", style = ~ style(color = "black",font.weight = "bold")),
                         `% change` = formatter("span", style = x ~ style(color = ifelse(x < 0, "green", ifelse(x > 0, "red", "black")))),
                         `Town rank` = formatter("span", 
                                       style = ~ style(color = ifelse(`townrank2` < `townrank1`, "green", ifelse(`townrank2`>`townrank1`,"red",""))),
                                               ~ icontext(ifelse(`townrank2` < `townrank1`, "arrow-up", ifelse(`townrank2`>`townrank1`,"arrow-down","")), `townrank1`)),
                         area_name=FALSE,
                         poi=FALSE,
                         rank=FALSE,
                         geometry=FALSE,
                         geoid=FALSE,
                         STATEFP=FALSE,
                         COUNTYFP=FALSE,
                         TRACTCE=FALSE,
                         BLKGRPCE=FALSE,
                         AFFGEOID=FALSE,
                         GEOID=FALSE,
                         NAME=FALSE,
                         LSAD=FALSE,
                         ALAND=FALSE,
                         AWATER=FALSE,
                         change=FALSE,
                         prob_sum_home=FALSE,
                         prob_sum_contact_lastweek2=FALSE,
                         prob_sum_contact_lastweek=FALSE,
                         townrank1 = FALSE,
                         townrank2 = FALSE,
                         pchange = FALSE
                         )) ,
                         rownames=FALSE,
                         class='cell-border stripe'
)
```

```{r child = 'about_report.md'}
```
```{r child = 'ack.md'}
```
```{r child = 'contact.md'}
```




